### 2018-03-19 submission/manuscript.Rmd
spell check

A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, PCR, and sequencing steps. Generally, not using a preservation media and leaving samples at room temperature has been shown to cause overgrowth of low abundance members of the fecal bacterial community [@Amir2017]. Similarily, this overgrowth can still occur if the preservation media does not adequately inhibit growth [@preservation_Luo_2016; @preservation_Song_2016]. Reports have also shown that changes in specific community members might occur due to differing susceptibility to freeze thaw cycles amongst microbes [@Gorzelak2015]. Additionally, reagent contamination has been shown to add community members and the contribution of these contaminant members grows larger with lower biomass samples [@contamination_Salter2014]. Recent studies have shown that the majority of these biases due to either preservation or extraction tend to be smaller than the overall biological signal being measured [@preservation_Song_2016; @storage_Bassis_2017]. However, the contribution of PCR bias to this overall workflow is not well characterized since these studies use the same PCR approach while varying preservation or extraction method.

@@

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences in the median average per base error varied by HiFi DNA polymerase without a clear pattern across sub-sampling depth [Table S5]. The highest median average per base error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. The differences in the median average per base error rate between the different HiFi DNA polymerases was cycle dependent with Platinum having the largest changes versus other HiFi DNA polymerases [Figure 4B-C and Table S6]. The total sequences with at least one error was also cycle number dependent and differences between HiFi DNA polymerases could be drastically reduced by the use of the 'pre.cluster' step [Figure S1]. These differences in sequences with at least one error were mostly due to differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerases [Figure S1 & Table S7 & S8]. Finally, we did not observe a HiFi DNA polymerase dependent difference on base substitution rate [Figure S2]. Although sequence error is dependent on HiFi DNA polymerase some of these error dependent differences can be corrected using exsiting bioinformatic approaches.

>>>>

A typical 16S rRNA gene sequencing workflow can be divided into preservation, extraction, PCR, and sequencing steps. Generally, not using a preservation media and leaving samples at room temperature has been shown to cause overgrowth of low abundance members of the fecal bacterial community [@Amir2017]. Similarly, this overgrowth can still occur if the preservation media does not adequately inhibit growth [@preservation_Luo_2016; @preservation_Song_2016]. Reports have also shown that changes in specific community members might occur due to differing susceptibility to freeze thaw cycles amongst microbes [@Gorzelak2015]. Additionally, reagent contamination has been shown to add community members and the contribution of these contaminant members grows larger with lower biomass samples [@contamination_Salter2014]. Recent studies have shown that the majority of these biases due to either preservation or extraction tend to be smaller than the overall biological signal being measured [@preservation_Song_2016; @storage_Bassis_2017]. However, the contribution of PCR bias to this overall workflow is not well characterized since these studies use the same PCR approach while varying preservation or extraction method.

@@

***Sequence Error is Dependent on both HiFi DNA Polymerase and Cycle Number:***
Differences in the median average per base error varied by HiFi DNA polymerase without a clear pattern across sub-sampling depth [Table S5]. The highest median average per base error rates were for the Kappa HiFi DNA polymerase [Figure 4]. This error rate was minimally affected by both the 'pre.cluster' step and chimera removal by VSEARCH [Figure 4]. The differences in the median average per base error rate between the different HiFi DNA polymerases was cycle dependent with Platinum having the largest changes versus other HiFi DNA polymerases [Figure 4B-C and Table S6]. The total sequences with at least one error was also cycle number dependent and differences between HiFi DNA polymerases could be drastically reduced by the use of the 'pre.cluster' step [Figure S1]. These differences in sequences with at least one error were mostly due to differences in Accuprime^TM^ and Platinum versus the other HiFi DNA polymerases [Figure S1 & Table S7 & S8]. Finally, we did not observe a HiFi DNA polymerase dependent difference on base substitution rate [Figure S2]. Although sequence error is dependent on HiFi DNA polymerase some of these error dependent differences can be corrected using existing bioinformatic approaches.

### 2018-03-19 submission/manuscript.tex
spell check

A typical 16S rRNA gene sequencing workflow can be divided into
preservation, extraction, PCR, and sequencing steps. Generally, not
using a preservation media and leaving samples at room temperature has
been shown to cause overgrowth of low abundance members of the fecal
bacterial community (Amir et al., 2017). Similarily, this overgrowth can
still occur if the preservation media does not adequately inhibit growth
(Song et al., 2016; Luo et al., 2016). Reports have also shown that
changes in specific community members might occur due to differing
susceptibility to freeze thaw cycles amongst microbes (Gorzelak et al.,
2015). Additionally, reagent contamination has been shown to add
community members and the contribution of these contaminant members
grows larger with lower biomass samples (Salter et al., 2014). Recent
studies have shown that the majority of these biases due to either
preservation or extraction tend to be smaller than the overall
biological signal being measured (Song et al., 2016; Bassis et al.,
2017). However, the contribution of PCR bias to this overall workflow is
not well characterized since these studies use the same PCR approach
while varying preservation or extraction method.

@@

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences in the median average per base error
varied by HiFi DNA polymerase without a clear pattern across
sub-sampling depth {[}Table S5{]}. The highest median average per base
error rates were for the Kappa HiFi DNA polymerase {[}Figure 4{]}. This
error rate was minimally affected by both the `pre.cluster' step and
chimera removal by VSEARCH {[}Figure 4{]}. The differences in the median
average per base error rate between the different HiFi DNA polymerases
was cycle dependent with Platinum having the largest changes versus
other HiFi DNA polymerases {[}Figure 4B-C and Table S6{]}. The total
sequences with at least one error was also cycle number dependent and
differences between HiFi DNA polymerases could be drastically reduced by
the use of the `pre.cluster' step {[}Figure S1{]}. These differences in
sequences with at least one error were mostly due to differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerases {[}Figure S1 \& Table S7 \& S8{]}. Finally, we did not
observe a HiFi DNA polymerase dependent difference on base substitution
rate {[}Figure S2{]}. Although sequence error is dependent on HiFi DNA
polymerase some of these error dependent differences can be corrected
using exsiting bioinformatic approaches.

>>>>

A typical 16S rRNA gene sequencing workflow can be divided into
preservation, extraction, PCR, and sequencing steps. Generally, not
using a preservation media and leaving samples at room temperature has
been shown to cause overgrowth of low abundance members of the fecal
bacterial community (Amir et al., 2017). Similarly, this overgrowth can
still occur if the preservation media does not adequately inhibit growth
(Song et al., 2016; Luo et al., 2016). Reports have also shown that
changes in specific community members might occur due to differing
susceptibility to freeze thaw cycles amongst microbes (Gorzelak et al.,
2015). Additionally, reagent contamination has been shown to add
community members and the contribution of these contaminant members
grows larger with lower biomass samples (Salter et al., 2014). Recent
studies have shown that the majority of these biases due to either
preservation or extraction tend to be smaller than the overall
biological signal being measured (Song et al., 2016; Bassis et al.,
2017). However, the contribution of PCR bias to this overall workflow is
not well characterized since these studies use the same PCR approach
while varying preservation or extraction method.

@@

\textbf{\emph{Sequence Error is Dependent on both HiFi DNA Polymerase
and Cycle Number:}} Differences in the median average per base error
varied by HiFi DNA polymerase without a clear pattern across
sub-sampling depth {[}Table S5{]}. The highest median average per base
error rates were for the Kappa HiFi DNA polymerase {[}Figure 4{]}. This
error rate was minimally affected by both the `pre.cluster' step and
chimera removal by VSEARCH {[}Figure 4{]}. The differences in the median
average per base error rate between the different HiFi DNA polymerases
was cycle dependent with Platinum having the largest changes versus
other HiFi DNA polymerases {[}Figure 4B-C and Table S6{]}. The total
sequences with at least one error was also cycle number dependent and
differences between HiFi DNA polymerases could be drastically reduced by
the use of the `pre.cluster' step {[}Figure S1{]}. These differences in
sequences with at least one error were mostly due to differences in
Accuprime\textsuperscript{TM} and Platinum versus the other HiFi DNA
polymerases {[}Figure S1 \& Table S7 \& S8{]}. Finally, we did not
observe a HiFi DNA polymerase dependent difference on base substitution
rate {[}Figure S2{]}. Although sequence error is dependent on HiFi DNA
polymerase some of these error dependent differences can be corrected
using existing bioinformatic approaches.
